---
title: "Psychometrics (class) scale development"
subtitle: "Do you have a need for control in social situations?"
bibliography: sncss.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
---

This is bit of a write up for my graduate class in psychometrics.  I thought it would be fun to show the process by which my class _designed_ and _tested_ a scale.  I use those accents because this was a class activity and we completed this in a summer semester.  This isn't a valid measure, so please don't use it.

## Introduction

Controlling behaviors have the potential to impact group activities and projects in a disruptive manner.  Natural leadership may be subverted by another actor’s desire to maintain power over the group for any number of reasons.  Determining an estimate of an individual’s control would be beneficial in identifying those who may have the potential to cause negative outcomes and proactively remedying or preventing an issue from manifesting.  

Motivations for behind seeking control can come from different initial factors.  Burger and Cooper compiled a small set of theoretical approaches and statements understanding the reason for seeking control and power [-@burger1979desirability].  An individual may be motivated by wanting to demonstrate their own competence in their life.  Another may be motivated because of power or ability.  Control can also be sought as means to end stress and taking autonomy of one’s life.  

People who display a strong sense of control are not always viewed favorably by those they wish to act upon.  Suggestions of personal gain or a superiority complex are easily expressed when expertise or entitlement is not readily known.  Some leaders fall into the role while others actively seek it.  More accepted forms of leadership do not build upon an open foundation of manipulation but by making sacrifices that are for the best of the group.  Unselfish acts may be modulated by the leader’s own empathy and the consideration for each other individual affected by the decision.  Threats to leadership can be provoke protective responses.  When an alarming situation arises, acts of aggression could follow to preserve status.  An uninvolved role would not be as beneficial if action is required.

We have set out to create a scale to measure the degree of control an individual possesses in relation to social situations and to tests its internal reliability and validation against like, unlike, and unrelated measures.

### Scale Testing

## Methods

### Participants

Graduate students currently enrolled in a graduate level psychometrics class at West Chester University of Pennsylvania (_n_ = 11) volunteer to complete the assessment.  Additional subjects were recruited by the students (_n_ = 8) to increase the sample size.  Participants filled out scantron sheets with the unidentified scales.  Order effects were not accounted for.  No demographic information was obtained to maintain anonymity with a small sample size and convenience in sampling.

### Instruments

#### Scale for the Need of Control in Social Settings (SNCSS).

The SNCSS scale consists of 31 items with a maximum score of 155 [see @Appendix].  Each item is scored on a Likert scale from Strongly Disagree to Strongly Agree with reverse scoring on select items.  Items were developed after an in-class discussion of defining the construct of control.  A prior literary search was not made before this intuition driven discussion.  In initially 50 items were developed and 19 removed due either not maintaining a strong face validity for the construct of control or redundancy.  All students in the graduate level psychometrics class contributed to refining items and removing items with the aid of the professor.

#### Adaptive and Aggressive Assertiveness Scale (AAA-S)

The AAA-S [@thompson2011adaptive] consists of 19 situations with a possible response or responses which are rated from a 5-point scale of _Never_ to _Always_, with non-labelled intermediate values, for the extent to which the response best describes how the test taker would response.  The scale has a total of 30 responses posed; 15 measuring degree of Aggressive Assertiveness and 15 measuring degree of Adaptive Assertiveness.  In a sample of 261 psychology students, the developers reported α coefficient value of .82 and Spearman-Brown reliability of .79 for Adaptive assertive; and an α coefficient of .87 and Spearman-Brown reliability of .71 for Aggressive assertiveness.  

#### The Basic Empathy Scale in Adults (BES-A)

The BES-A [@carre2013basic] is a 20-item measure to assess empathy specifically in adults developed from the BES for children and adolescence.  The scale consists of an 11-item sub scale for Affective empathy and a 9-item sub scale for Cognitive empathy.  Each item is answered on a 5-point scale Likert scale.  With a sample of 370 participants drawn from psychology and social science students, employees, and retired persons, the scale authors reported a Cronbach α of .84 and reliability coefficient of .637 for the Affective empathy subscale; and a Cronbach α of .71 and reliability coefficient of .373 for the Cognitive empathy subscale.

#### Leadership Self-Report Scale (LSRS)

The LLS [@dussault2013leadership] is a 46-item questionnaire to measure leadership factors.  The scale consists of the following factors:  Transformational leadership (25 items); Transactional leadership (15 items); and Laissez-faire (6 items).  The Transformational leadership factor is composing of three scales: Charisma factor (8-items); Intellectual stimulation (8-items); and individual stimulation (8 items).  The Transactional leadership factor is composed of two scales:  Contingent reward (8 items); and Management-by-exception (7 items).  Each item is rated on a 4-point scale from Strongly Disagree to Strongly Agree.  Higher scores on the total scale and each subscale suggest higher levels of leadership.  Dussault and colleagues reported a Cronbach α reliability of .85 for the Transformational leadership scale, .78 for the Transactional leadership scale, and .65 for the Laissez-faire scale, with a sample of 222 participants from Quebec French schools.

#### The Narcissism Personality Inventory
NPI-16 [@ames2006npi] is an abbreviated version of the original 40 item Narcissism Personality Inventory.  Each item proses a pair of statements to choose between based on which statement best aligns with the test taker’s feelings and beliefs about themselves.  In the validation study with a sample of 776 undergraduate psychology students, the developers reported an alpha of .72 with the NPI-16 and an alpha of .84 with the full length 40-item NPI.  The NPI-16 was correlated with the unused 24 items of the NPI-40 at _r_ = .71.

## Results & Data Analysis

### Data Wrangling

```{r load_libraries, comment = "", message = F, warning = F}
library(psychometric)
library(reshape2)
library(broom)
library(tidyverse)
library(kableExtra)
select <- dplyr::select
```

```{r view_data, comment = ""}
dat <- read.csv("sncss.csv")
glimpse(dat)
```

We're also going to want to make a separate table for our total scores.

```{r create_totals, comment = ""}
tots <- dat %>%
  group_by(id, scale, domain) %>%
  summarise(item = "total",
            total = sum(response))
glimpse(tots)
```

And lastly we want to use the `broom` package to create a **tibble** which contains some item analysis

```{r view_tibble, comment = "", warnings = F}
ie_tib <- dat %>%
  group_by(name) %>%
  do(ie = dcast(., id ~ item, value.var = "response") %>%
       select(-id) %>%
       item.exam() %>%
       mutate(id = rownames(.))) %>%
  unnest() %>%
  select(name, id, everything())
kableExtra::kable(ie_tib) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```


### Reliability

An item analysis was performed on the 31 original items using the `psych` and `psychometric` packages.  The analysis yielded a Cronbach’s α of .735.  A threshold for item reliability of .250 was set to determining whether an item would be removed or not.  Of the 31 items, 14 yielded reliability coefficients below this threshold and were subsequently removed.  A second item analysis was performed with the remaining 17 items.  Although item 30 in the second analysis showed a correlation coefficient of r = .182, removal of the item would have had only a slight positive effect on Cronbach’s α; an increase from .808 to .814 if deleted.  The question itself, “_I like to micromanage others_” was agreed to have a strong face value and retained for the SNCSS-17.

```{r cronbach alpha, comment = ""} 
df <- dat %>%
  filter(name == "CON_full") %>%
  dcast(id ~ item, value.var = "response") %>%
  select(-id)

cronbach(df)$`Alpha`

keep <- ie_tib %>%
  filter(name == "CON_full") %>%
  filter(Item.Tot.woi > .250) %>%
  select(id) %>%
  unlist() %>%
  as.numeric()

cronbach(df[, keep])$`Alpha`
```

### Validity

Internal consistency coefficients were computed for each factor/subscale: (BES-A) Affective Empathy; (BES-A) Cognitive Empathy; (AAA-S) Adaptive Assertiveness; AAA-S Aggressive Assertiveness; NPI.  As the sample size was limited, a more generous α = .15 was decided for measuring statistical significance.  The results of these analyses showed only one large difference in alpha coefficients.  The Laissez-faire subscale of the Leadership Self-Report Scale was originally reported to have a coefficient .65 by the developers but was found to have a much lower .352 in the present study.  The Laissez-faire subscale’s low correlation and significance (_r_ = -.002, _p_ = .993) is not surprising given its lowered alpha coefficient, low number of items, and small sample.

```{r validity, comment = ""}
df <- dat %>%
  group_by(name) %>%
  do(cronbach = dcast(., id ~ item, value.var = "response") %>%
       select(-id) %>%
       cronbach() %>%
       data.frame()) %>%
  unnest()

kableExtra::kable(df, format = "markdown")
```

#### Items dropped

- **Item 4** of the BES-A Affective Empathy scale “_I get frightened when I watch characters in a good scary movie_” was removed as it was deleted from the original validation because of its weakness in a two-factor and three-factor model [@carre2013basic].
- The following items from the AAA-S were removed from this analysis as they were also removed from the original study due to low factor loadings – values below .3: **4a**, **4b**, **7b**, **9a**, **13b**, and **16b**.
- Item **11a** was removed as an item 11b was noted in the original article but not on the scale
- **14a** was removed for a similar reason as well
- Items **13a** (“_If I am at a performance and someone keeps talking loudly, I … Would tell the person to shut up._”) and **16a** (“_If someone cuts in line ahead of me at the moves, I … Start making loud comments about how rude the person is._”) were both labeled as adaptive responses which appeared incorrect.  As it was not easily determined whether there was an error in numbering or labelling, these items were removed.

#### Convergent Validity

The AAA-S Aggressive assertiveness and NPI-16 were chosen to have expected positive correlations with the SNCSS-17 on the proposed concept that individuals more likely to need or assume control would have greater or more frequency aggressive behaviors and greater views of themselves.  Table 2 summarized these results.  Low and non-statistically significant correlations were found for Aggressive assertiveness and the NPI-16 (_r_ = .289, _p_ = .230; _r_ = .271, _p_ = .262).

The BES-A Affective empathy and Cognitive empathy, AAA-S Adaptive assertiveness, and Leadership Self-Report Scale Laissez-faire were expected to negatively correlate with the SNCSS-17.  A necessity for control was originally viewed as requiring less empathetic considerations.  Behaviors and attitudes underlining a desire for control was not recognized as a demonstration of an adaptive or uninvolved leader but more so a destructive and demanding one.  A significant and strong positive correlation was found for the Adaptive assertiveness (_r_ = .493, _p_ = .032) but low correlation and significance were found for Affective empathy (_r_ = .325, _p_ = .174), Cognitive empathy (_r_ = .297, _p_ = .174), and Laissez-faire (_r_ = -.002, _p_ = .993).

```{r, comment = ""}
## To add in CON_17
df <- dat %>%
  filter(name == "CON_full") %>%
  dcast(id ~ item, value.var = "response") %>%
  select(-id) %>%
  rowSums()

table <- dat %>%
  filter(name != "CON_full") %>%
  group_by(name) %>%
  do(x = dcast(., id ~ item, value.var = "response") %>%
       select(-id) %>%
       rowSums() %>%
       cor.test(., df) %>%
       tidy()) %>%
  unnest()
kableExtra::kable(table, format = "markdown") 
```

Affective empathy and cognitive empathy had surprisingly positive correlation coefficients, thus contradicting the initial stance that attempting to acquire control is inversely correlated with an individual’s potential empathy.  The thoughts that all desire for control stems from selfish reasonings would need to reexamine.  Parsing out motivations and desires behind control, other than personal power, would help in understanding this trend in the analysis.

Laissez-faire leadership subscale only contained 6 items, and with a sample size of _n_ = 17, the underwhelming results are of no surprise. The correlation and significance were negligible; thus no substantial considerations can be made.

The strong positive correlation of the SNCSS-17 to the AAA-S Adaptive assertiveness subscale was an initial surprise.  A strong desire for control was not originally viewed as an adaptive trait in leadership.  However, adaption could function as a means of maintaining control.  Miron and Brehm [-@miron2006reactance] explain _reactance_ as the emotional-motivational response to situations perceived to constrict choice.  The individual reacts to these events with increased efforts to demonstrate their own control.  An adaptive individual would better be able to maintain their control in different situations than a non-adaptive or aggressive leader.  Take for instance item 4b and 12b of the AAA-S, “_I am at the grocery store and several of my items ring up incorrectly, I … Ask the cashier to do a price check on the particular items_” and “_If someone I don’t know well disagrees with me during a conversation, I … Continue elaborating on my opinion until the person understands it._”.   Both items demonstrate an adaptive reactance to an event which may be more successful than their alternatives; getting angry.  Alternative means of a solution (_e.g._, the item price-check) and persistence or persuasion, would do well for an individually trying to maintain control in a situation.

#### Discriminant validity

Two subscales from the Leadership Self-Report Scale, Transformational and Transactional leadership, were not hypothesized to have any indication for control.  A weak and non-significant correlation was found for Transactional leadership (_r_ = .026, _p_ = .915).  However, a significant negative correlation was found for Transformational leadership (_r_ = -.356, _p_ = .135).

The results of Transactional leadership subscale confirm the intuition that there would exist no correlation between this measure and the SCNSS-17.  However, the correlation between Transformational leadership and desire for control require some explanation.  Transformational leadership.  This measure is based upon self-reports of cheerfully, optimism, and encouragement.  These are inspirational factors from a leader.  Glass and Singer [-@glass1972behavioral] demonstrate that when faced with pestering, uncontrollable stimuli, an individual’s cognitive ability is impaired and stress is elevated.  Perhaps our individuals that desire control are negatively affected by the lack of it or are perpetually worried about losing their control.  Their stress, from maintaining their position, may factor into their overall expression of joyfulness and transformational values.

```{r corr_plots}
dat %>% 
  dcast(id ~ name, value.var = "response",fun.aggregate = sum) %>%
  melt(id = c("id", "CON_full")) %>%
  ggplot(aes(x = CON_full, y = value)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~variable, scales = "free") +
  labs(y = NULL)
```

### Factor analysis

A two component factor analysis was performed on the SCNSS-17 with a oblimin rotation using the `gpa` package.  See Table 3 for a summary.  Results demonstrated a strong factor loading of 5 items into one component and 8 items into a second component.  Four items did not show strong factor loading values for one particular component.  That is, the factor loading values for these components were below .30 for either component or showed cross loading between the two factors [@costello2005best].  After analyzing each item in relation to these components, the sets were categorized into **Motivation** items and **Ability** items.

The factor loading analysis showed a suspicion that the items deemed suitable from class discussion and then filtered out with an item analysis, may be the cumulation of more than one factor in a need for control.  After the items were categorized by component, a determination of these components was made.  Ability items (_e.g._, “_It feels natural for me to lead conversations when working on group projects._”, “_I can get my way through talking to people._”) illustrate the possible magnitude of control an individual may posses through either persuasion or a feeling of nature leadership.  Items from the second component, Motivation (_e.g._, “_When I go on a trip, I like to keep a predetermined schedule._”, “_When things don’t go my way, I get upset._”, “_Typically, my judgement is better than my peers_”) illustrate desires to be involved in planning and making decisions as well reactions to possible threats of losing or not gaining control.  Separate analyses had not been performed with these possible subscales.  

```{r fa, comment = "", message = F, warning = F}
df <- dat %>%
  filter(name == "CON_full") %>%
  dcast(id ~ item, value.var = "response") %>%
  select(-id)
fa <- psych::fa(df, nfactors = 2, rotate = "varimax")
fa$loadings
plot(fa)
```

## Other considerations

Burger and Cooper [-@burger1979desirability] developed a 20-item Desirability of Control Scale (DC) which measures an individual’s general desire for control over events.  The items were specifically created to related to a direst for control in relation to events in an individual’s environment in a general and specific terms.  Each statement is rated on a 7-point scale from “_This statement doesn’t apply to me at all_” to “_This statement always applies to me_”.  The DC was intended to measure Assertion, Decision, and Activity in as factors of desire.  The DC has many similar items to the SNCSS-17 and would appear to be a great resource and to determine convergent validity but was unused in the present study as it was obtained only after the analyses were completed.  

Overall, the results from the analysis with a limited sample suggest some relevance to the current measure.  Items could be modified further or account for more slight redundancies and gearing towards a bifactor consideration for actions involving a leadership role and attitudes and motivations for seeking control.  Another relevant scale was identified that could supply greater backing for the validity of the current measure and should be explored. 

## Appendix

## References
